Using Scipion for Stream Image Processing at EM Facilities

### 2018-03-21 1.introduction.tex
casa

>>>>

The organization of this document is as follows. First, we present the modifications introduced in \scipion to adapt it to large centralized microscopy facilities. Then we present a typical workflow to demostrate \scipion capabilities. Finally,  we describe in detail how \scipion is being used in three representative facilities.

### 2018-03-21 2.overall.tex
casa

\subsection{Stream Processing}
The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as it is produced or received.
In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols. For example, as movies are imported to the system they are aligned and their CTF is estimated. In \scipion version 1.1, users may find  the first steps in the image processing adapted to streaming, including movie alignment and CTF estimation. We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2. In the developers version, several 2D classification algorithms are being adapted for this new way of processing.

%\scipion is a multilayered application, the first layer handles the interaction with the user (\emph{presentation layer}), the second saves and retrieves the data used by the application  (\emph{data layer}) and finally, the third layer (\emph{business layer}) process the data following the user instructions.  
\scipion stream implementation make extensive use of threads. %A thread is the smallest unit of processing that can be scheduled by an operating system. 
Using threads, a programmer can effectively create two or more tasks that run at the same time and share the protocol resources (i.e. memory). 
%is based on periodic updates of the protocol input and output data during the image processing. 
%Each protocol execution in \scipion is an independent process that can create processing threads. 
Each protocol that works in streaming mode has one of its threads monitoring input and updating output  while other threads take care of the data processing. This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several computations are executed  concurrently, instead of sequentially. Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
%In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line. %Moreover, it also allows to quickly develop new workflows .

@@

We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance. That is, \scipion provides detailed information about the origin of the data XXXXX

 easily share traceable information and research results is central to their use.

in such a way  ensuring reproducibility in scientific research

that data will be traceable and reproducible

keeps a historical record of how the data was generated reproducibility and traceability.

@@

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB.

@@

As commented in section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols. By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed. \protocol{System monitor} reports on  CPU and GPU utilization, memory consumption and, disk I/O. \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range. Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server so this information can be accessed by users who are not physically in the microscope room.

>>>>

\subsection{Stream Processing}
The main change introduced in \scipion is \emph{stream processing}, that is, computing on data directly as it is produced or received.
In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols. For example, as movies are imported to the system they are aligned and their CTF is estimated. In \scipion version 1.1, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment and CTF estimation. We have also implemented particle picking and extraction, that are going to be released soon in the new version 1.2. In the developers version, several 2D classification algorithms are being adapted for this new way of processing.

%\scipion is a multilayered application, the first layer handles the interaction with the user (\emph{presentation layer}), the second saves and retrieves the data used by the application  (\emph{data layer}) and finally, the third layer (\emph{business layer}) process the data following the user instructions.  
\scipion stream implementation make extensive use of threads. %A thread is the smallest unit of processing that can be scheduled by an operating system. 
Using threads, a programmer can effectively create two or more tasks that run at the same time and share the protocol resources (i.e. memory). 
%is based on periodic updates of the protocol input and output data during the image processing. 
%Each protocol execution in \scipion is an independent process that can create processing threads. 
Each protocol that works in streaming mode, has one of its threads monitoring input and updating output  data while other threads take care of the data processing. This clear separation between data access and data processing marks an important difference between running a script or using \scipion since it facilitates a form of computing in which several jobs are executed  concurrently, instead of sequentially. Moreover, \scipion workflow engine, allows users to modify the workflow on the fly without modifying a single code line.
%In practice, the latter follows  a more general and extensible approach that allows users to modify the workflow on the fly without modifying a single code line. %Moreover, it also allows to quickly develop new workflows .

@@

We would like to finish this section assuring that despite the modifications introduced to \scipion data and process model, the framework provides full data provenance. That is, \scipion provides detailed information about the origin of the data  and the process to generate it.

@@

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB cryoEM Service.

@@

As commented in section \ref{overall}, to monitor, chart and create alerts  we have developed a set of metaprotocols. By default \protocol{system}, \protocol{CTF} and \protocol{report monitors} are executed. \protocol{System monitor} reports on  CPU and GPU usage, memory consumption and, disk I/O. \protocol{CTF monitor} plots the CTF defocus and astigmatism  and it will raise an alarm if these parameters are not within a given range. Finally the \protocol{report} protocol pushes all the created plots and summaries to a web server so this information can be accessed by users who are not physically in the microscope room.

### 2018-03-21 3.workflow.tex
casa

So far, \scipion is being used routinely in four Cryo-EM facilities: \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}),  \emph{European Synchrotron Radiation Facility} (ESRF, \url{http://www.esrf.eu/}), \emph{Diamond Light Source Synchrotron} (Diamond, \url{http://www.diamond.ac.uk}) and \emph{National Center for Biotechnology} (CNB, \url{http://www.cnb.csic.es}). \scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed. There, \scipion was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.

%\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipion has been installed.

>>>>

So far, \scipion is being used routinely in seven Cryo-EM facilities:
\begin{itemize}
 \itemsep0em

 \item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
 \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se})
 \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
 \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html}) and
 \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
 \item \emph{Molecular Microscopy Consortium} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
 \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]

\end{itemize}

\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed. There, \scipion was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.

SAY WE ARE GOING TO DESCRIBE THREE REPRESENTAIVE CENTER
%\scipion is a very flexible framework that allows to create many different workflows and to choose among several algorithms at each workflow-step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since different use cases may clarify \scipionbox capabilities, in the following we describe the pipeline offered by default to the users of three different facilities where \scipion has been installed.
